---
title: "FDR Plot Redone"
author: "Brendan Law"
date: "18/05/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a repeat of `real_data_analysis.Rmd` to get a plot that's more like Figure 1b on Shim and Stephens' original Wavelet paper here: https://github.com/heejungshim/WaveQTL/blob/master/doc/paper/Shim_2014.pdf

This takes the data from `real_data_analysis.Rmd` and tries to do the same sort of bar plot.

Note that as of 2025, to install the qvalue package from Bioconductor, refer to this website: https://bioconductor.org/packages/release/bioc/html/qvalue.html

Current process is to run:
```
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("qvalue")
```

Data pre-processing functions:
```{r}
library(data.table)

hmt_pvals = readRDS("../data/real_data_analysis/hmt_pvals.RDS")
nohmt_pvals = readRDS("../data/real_data_analysis/nohmt_pvals.RDS")

hmt_pvals = hmt_pvals[!is.na(pval)]
nohmt_pvals = nohmt_pvals[!is.na(pval)]

pval = data.table(
  pval.hmt = as.numeric(hmt_pvals$pval)
  ,pval.nohmt = as.numeric(nohmt_pvals$pval)
)
```


```{r}
# check histograms
hist(pval$pval.hmt, breaks=100, main = "hmt")

hist(pval$pval.nohmt, breaks=100, main = "nohmt")

# apply qvalue package
library("qvalue")
qval.hmt = qvalue(pval$pval.hmt) # 1
qval.nohmt = qvalue(pval$pval.nohmt) # 0.002288085

# check the proportion of null cases
qval.hmt$pi0 # 0.8354974
qval.nohmt$pi0 # 0.8496002

# possible values of FDR
alpha.list = seq(0, 0.1, by=0.001)
length(alpha.list)
## 101

# count the number of significant tests at a given FDR
num.hmt = num.nohmt = rep(NA, length(alpha.list))
for(i in 1:length(alpha.list)){
  num.hmt[i] = sum(qval.hmt$qvalues < alpha.list[i])
  num.nohmt[i] = sum(qval.nohmt$qvalues < alpha.list[i])
}

# number of significant tests at FDR = 0.05
wh = which(alpha.list == 0.05)
num.hmt[wh] # 208
num.nohmt[wh] # 147
```


Make bar plot of number of significant vs non significant dsQTLs identified by HMT vs non HMT approach at FDR of 0.02, 0.05.

The whole idea here is to find:
- dsQTLs which are identified by both (significant in both); that's the dark green.
- dsQTLs identified in HMT but not Wavelet (light green for HMT)
- dsQTLs identified in Wavelet but not HMT (light green for Wavelet)

```{r}
# at 0.05, here are:
# significant dsQTLs under HMT
hmt_sig_005 = which(qval.hmt$qvalues < .05)
# not significant dsQTLs under HMT
hmt_not_sig_005 = which(qval.hmt$qvalues > .05)

# significant dsQTLs under wavelet
nohmt_sig_005 = which(qval.nohmt$qvalues < .05)
# not significant dsQTLs under wavelet
nohmt_not_sig_005 = which(qval.nohmt$qvalues > .05)

# dsQTLs significant in both
dsqtls_sig_both_005 <- intersect(hmt_sig_005, nohmt_sig_005)
length(dsqtls_sig_both_005)
# dsQTLs significant in HMT but not Wavelet
dsqtls_sig_hmt_only_005 <- setdiff(hmt_sig_005, nohmt_sig_005)
length(dsqtls_sig_hmt_only_005)
# dsQTLs significant in Wavelet but not HMT
dsqtls_sig_wavelet_only_005 <- setdiff(nohmt_sig_005, hmt_sig_005)
length(dsqtls_sig_wavelet_only_005)
```

```{r}
# at 0.02, here are:
# significant dsQTLs under HMT
hmt_sig_002 = which(qval.hmt$qvalues < .02)
# not significant dsQTLs under HMT
hmt_not_sig_002 = which(qval.hmt$qvalues > .02)

# significant dsQTLs under wavelet
nohmt_sig_002 = which(qval.nohmt$qvalues < .02)
# not significant dsQTLs under wavelet
nohmt_not_sig_002 = which(qval.nohmt$qvalues > .02)

# dsQTLs significant in both
dsqtls_sig_both_002 <- intersect(hmt_sig_002, nohmt_sig_002)
length(dsqtls_sig_both_002)
# dsQTLs significant in HMT but not Wavelet
dsqtls_sig_hmt_only_002 <- setdiff(hmt_sig_002, nohmt_sig_002)
length(dsqtls_sig_hmt_only_002)
# dsQTLs significant in Wavelet but not HMT
dsqtls_sig_wavelet_only_002 <- setdiff(nohmt_sig_002, hmt_sig_002)
length(dsqtls_sig_wavelet_only_002)
```


```{r}
# Make bar plots
library(ggplot2)
library(dplyr)


df <- data.table(
  fdr = rep(c("FDR = 2%","FDR = 5%"), each=2),
  method = rep(c("HMT","Wavelet"),2),
  num_sig_both = rep(c(length(dsqtls_sig_both_002), length(dsqtls_sig_both_005)), each=2),
  num_sig_this_only=c(
    length(dsqtls_sig_hmt_only_002),
    length(dsqtls_sig_wavelet_only_002),
    length(dsqtls_sig_hmt_only_005),
    length(dsqtls_sig_wavelet_only_005)
  )
)

df_melt <- melt.data.table(df,id.vars=c("fdr","method"))

ggplot(df_melt, aes(x = factor(method), y = value, group = factor(fdr), fill = variable)) +
  geom_col(position="stack") +
  facet_wrap(~fdr) +
  labs(x = "Method", y = "Number of significant tests", title = "") +
  theme_minimal() + 
  scale_fill_discrete(labels=c("Significant in the other approach", "Not significant in the other approach"), name="") +
  theme(legend.position = "bottom")
```

```{r}
# Ensure your data is in the right format
df_melt$method <- as.factor(df_melt$method)
df_melt$variable <- as.factor(df_melt$variable)
df_melt$fdr <- as.factor(df_melt$fdr)

# Unique FDR levels
fdr_levels <- levels(df_melt$fdr)
n_panels <- length(fdr_levels)

# Define fill colors
fill_colors <- c("#006400", "#00FF00")
legend_labels <- c("Significant in the other approach", "Not significant in the other approach")

path = "../figures/compare_significant_tests.png"
png(path, units="in", width = 7, height = 3, res =300)
# Create a 1-row, 3-column layout: 2 plots + 1 legend
layout(matrix(c(1, 2, 3), nrow = 1), widths = c(1, 1, 1))  # Third column is for the legend

# Set margins and outer margins
par(mar = c(5, 4, 2, 1), oma = c(1, 0, 1, 0))

# Compute shared y-axis max
max_y <- 250
# max_y <- 0
# for (i in seq_along(fdr_levels)) {
#   sub_data <- subset(df_melt, fdr == fdr_levels[i])
#   tab <- xtabs(value ~ variable + method, data = sub_data)
#   max_y <- max(max_y, sum(tab))
# }

# Plot each facet
for (i in seq_along(fdr_levels)) {
  sub_data <- subset(df_melt, fdr == fdr_levels[i])
  tab <- xtabs(value ~ variable + method, data = sub_data)
  
  barplot(
    tab,
    beside = FALSE,
    col = fill_colors,
    main = fdr_levels[i],
    xlab = "",
    ylab = if (i == 1) "Number of significant tests" else "",
    las = 0,
    ylim = c(0, max_y),
    yaxt = if (i == 1) "s" else "n"
  )
}

# Plot empty plot for legend
par(mar = c(0, 0, 0, 0))
plot.new()
legend("center",
       legend = legend_labels,
       fill = fill_colors,
       bty = "n",
       xpd = TRUE)
dev.off()

```