---
title: "Simulated data: auROC plots for various effect strengths and sizes - one instance"
author: "Brendan Law"
date: "09/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

These are plots which summarise the performance of WaveQTL vs WaveQTL_HMT, based on area under an receiver operating characteristic curve (AUC). These AUC plots summarise each algorithm's ability to detect signals on 200 simulated datasets. More info can be found in the manuscript's section on simulation.

Data pre-processing functions:
```{r}
extract_roc_plot_data <- function(sims_list){

  num_sims <- length(sims_list$null_waveqtl_lhood)
  preds_nohmt <- matrix(c(sims_list$null_waveqtl_lhood
                          ,sims_list$alt_waveqtl_lhood
                          ,rep(0,num_sims),rep(1,num_sims))
                        ,nrow = 2*num_sims,ncol = 2,byrow = F)
  preds_hmt <- matrix(c(sims_list$null_waveqtl_hmt_lhood
                        ,sims_list$alt_waveqtl_hmt_lhood
                        ,rep(0,num_sims),rep(1,num_sims))
                      ,nrow = 2*num_sims,ncol = 2,byrow = F)
  
  n = dim(preds_hmt)[1]
  
  O1 = order(preds_nohmt[,1], decreasing =TRUE)
  O2 = order(preds_hmt[,1], decreasing =TRUE)
  
  # True positives proportions or sensitivity
  C1  = c(0,cumsum(preds_nohmt[O1,2])) / sum(preds_nohmt[,2])
  C2  = c(0,cumsum(preds_hmt[O2,2])) / sum(preds_hmt[,2])
  
  # False positives proportions based on Model 1.
  FP1 = c(0,cumsum(1-preds_nohmt[O1,2])) / (n-sum(preds_nohmt[,2]))
  FP2 = c(0,cumsum(1-preds_hmt[O2,2])) / (n-sum(preds_hmt[,2]))
  
  return(list(
    FP1 = FP1
    ,C1 = C1
    ,FP2 = FP2
    ,C2 = C2
  ))
}
```

Process data:
```{r}
l8 <- readRDS("../data/auroc_simulation_data/one_instance/20220327_l8_paper_nogrp_od210_combined.RDS")
l16 <- readRDS("../data/auroc_simulation_data/one_instance/20220327_l16_paper_nogrp_od210_combined.RDS")
l24 <- readRDS("../data/auroc_simulation_data/one_instance/20220327_l24_paper_nogrp_od210_combined.RDS")
l8_roc_data <- lapply(l8,extract_roc_plot_data)
l16_roc_data <- lapply(l16,extract_roc_plot_data)
l24_roc_data <- lapply(l24,extract_roc_plot_data)
```

For each effect length, an auROC curves at various effect strengths for one instance of 200 simulated data points:
```{r}
cols.efsize = c("#104E8B","#FFA500","#006400")
lab.size = 1.2
y.lab.size = 1.2
x.lab.size = 1.2
line.thick = 1.2
pch.type = 19
cex.thick = 0.5
indices_to_use = c(1,3,4)
path = "../figures/auroc_by_eff_length.png"
png(path, units="in", width = 12, height = 4, res =300)
numfig = 3
nf <- layout(matrix(1:numfig,1,numfig,byrow=FALSE),TRUE)

# Effect length 8
par(mar = c(4,4,3,1))
plot(0,0, xlim = c(0,1), ylim = c(0,1), xlab = "False Positive Rate", ylab ="True Positive Rate", type="n", cex.lab=lab.size, axes=FALSE, main = "Effect Length 8")

col_idx <- 1
for(idx in indices_to_use){
  lines(l8_roc_data[[idx]]$FP1,l8_roc_data[[idx]]$C1, lty = 3, col=cols.efsize[col_idx], xaxt="n", yaxt="n", pch=pch.type, lwd=line.thick, cex=cex.thick)
  lines(l8_roc_data[[idx]]$FP2,l8_roc_data[[idx]]$C2, lty = 1, col=cols.efsize[col_idx], xaxt="n", yaxt="n", pch=pch.type, lwd=line.thick, cex=cex.thick)
  col_idx = col_idx + 1
}

lines(seq(0,1,length.out = 10000),seq(0,1,length.out = 10000), col="grey60", xaxt="n", yaxt="n", pch=pch.type, lwd=line.thick, cex=cex.thick)
axis(side = 2, cex.axis = y.lab.size)
axis(side = 1, cex.axis = x.lab.size)
legend(0.55,0.3
       ,legend = c("WaveQTL","WaveQTL-HMT",as.character(unlist(lapply(l8,function(x){x$params_list$effect_multiple})[indices_to_use])))
       ,lty = c(3,1,rep(1,length(indices_to_use)))
       ,col = c("black","black",cols.efsize)
       ,text.col = c("black","black",cols.efsize)
       ,cex = lab.size
       ,merge = FALSE, bg = "white", bty = "n", lwd = 1.5)
box()

# Effect length 16
par(mar = c(4,2.5,3,2.5))
plot(0,0, xlim = c(0,1), ylim = c(0,1), xlab = "False Positive Rate", ylab ="True Positive Rate", type="n", cex.lab=lab.size, axes=FALSE, main = "Effect Length 16")

col_idx <- 1
for(idx in indices_to_use){
  lines(l16_roc_data[[idx]]$FP1,l16_roc_data[[idx]]$C1, lty = 3, col=cols.efsize[col_idx], xaxt="n", yaxt="n", pch=pch.type, lwd=line.thick, cex=cex.thick)
  lines(l16_roc_data[[idx]]$FP2,l16_roc_data[[idx]]$C2, lty = 1, col=cols.efsize[col_idx], xaxt="n", yaxt="n", pch=pch.type, lwd=line.thick, cex=cex.thick)
  col_idx = col_idx + 1
}

lines(seq(0,1,length.out = 10000),seq(0,1,length.out = 10000), col="grey60", xaxt="n", yaxt="n", pch=pch.type, lwd=line.thick, cex=cex.thick)
axis(side = 2, cex.axis = y.lab.size)
axis(side = 1, cex.axis = x.lab.size)
box()

# Effect length 24
par(mar = c(4,1,3,4))
plot(0,0, xlim = c(0,1), ylim = c(0,1), xlab = "False Positive Rate", ylab ="True Positive Rate", type="n", cex.lab=lab.size, axes=FALSE, main = "Effect Length 24")

col_idx <- 1
for(idx in indices_to_use){
  lines(l24_roc_data[[idx]]$FP1,l24_roc_data[[idx]]$C1, lty = 3, col=cols.efsize[col_idx], xaxt="n", yaxt="n", pch=pch.type, lwd=line.thick, cex=cex.thick)
  lines(l24_roc_data[[idx]]$FP2,l24_roc_data[[idx]]$C2, lty = 1, col=cols.efsize[col_idx], xaxt="n", yaxt="n", pch=pch.type, lwd=line.thick, cex=cex.thick)
  col_idx = col_idx + 1
}

lines(seq(0,1,length.out = 10000),seq(0,1,length.out = 10000), col="grey60", xaxt="n", yaxt="n", pch=pch.type, lwd=line.thick, cex=cex.thick)
axis(side = 2, cex.axis = y.lab.size)
axis(side = 1, cex.axis = x.lab.size)
box()
```
